Large-scale harmonic patterns clustering

### 2010-03-11 Paper/Paper.tex
improvements

\subsection{Codebook Size in Function of Pattern Length}
We experiment on the randomness of the \textit{Shingles}, assuming
that there is some correlation between beats and bars.
The intuition is as follow: if the signal was totally random, every
time the signal size double, the codebook size must be squared.
To illustrate this, imagine a binary array of size $1$. We need two
codes to encode perfectly: $(0)$ and $(1)$. If the binary is now of
size $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.

@@

\subsection{Visualization}
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. We use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.

@@

In all our experiments we use Matlab's implementation of NMF:
nmflib\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.

\subsection{Training on our Codebook}
NMF provides us with a way of measuring how much information is retained
in our codebook. More specifically, we can train a codebook of size $500$
using our online VQ algorithm, then decompose that code into $10$ basis
using NMF. We can compare that to $10$ basis from training NMF on the
whole data.

@@

\section{Conclusion}
We are awesome. Here's how:
\begin{itemize}
\item large-scale
\item smarter shingles, based on beats and bars
\item clusters good for encoding
\item clusters good for something else?
\item free online API, code available\footnote{code not released yet to preserve 
submission's anonymity}
\item we have lab t-shirts
\end{itemize}

>>>>

\subsection{Codebook Size in Function of Pattern Length}
We experiment on the randomness of the \textit{Shingles}, assuming
that there is some correlation between beats and bars.
The intuition is as follow: if the signal was totally random, every
time the signal size double, the codebook size must be squared.
To illustrate this, imagine a binary array of length $1$. We need two
codes to encode perfectly: $(0)$ and $(1)$. If the binary array is now of
length $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.

@@

\subsection{Visualization}
Interpreting a codebook of size larger than $100$ is not obvious.
We can use an algorithm like locally linear embedding \cite{Roweis2000} (LLE) 
or multidimensional scaling \cite{Kruskal1964} to visualize the data in
2 dimensions. We use the former\footnote{implementation: 
\url{http://www.astro.washington.edu/users/vanderplas/coding/LLE/}}, the
result can be seen in Figure \ref{fig:lle}.

@@

In all our experiments we use the following NMF
implementation\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.

\subsection{Training NMF on our Codebook}
NMF provides us with a way of measuring how much information is retained
in our codebook. More specifically, we can train a codebook of size $500$
using our online VQ algorithm, then decompose that code into $10$ basis
using NMF. We can compare that to $10$ basis from training NMF on the
whole data.

@@

\section{Conclusion}
We presented a practical method to perform large-scale clustering of
harmonic patterns. We assessed the basic properties of the method through
experiments on a large collection of music. We suggested many ways
to interpret the data through visualization (LLE) and further encoding
(NMF). We discussed the possibility to move to even larger scales
and we provided our source code\footnote{Code not yet released to preserve
submission's anonymity}.
